GeneSet_Expansion

### 2020-04-22 code/GeneSet_Expansion.Rmd
update

Regions[ c("TREM2", "CD2AP", "MTCH2", "EED", "APOC1"), ]$V2<-Regions[ c("TREM2", "CD2AP", "MTCH2", "EED", "APOC1"), ]$V2+500000
Regions[ c("TREM2", "CD2AP", "MTCH2", "EED", "APOC1"), ]$V3<-Regions[ c("TREM2", "CD2AP", "MTCH2", "EED", "APOC1"), ]$V3-500000

@@

Candidates <- list()
for( name in as.character(Regions$V4) ){
  filterlist <- paste0( Regions[ name, ]$V1, ':', Regions[ name, ]$V2, ':', Regions[ name, ]$V3 )

@@

mark <- Sys.time()
for( name in as.character(Regions$V4) ){
 #for( Tissue in names(ExpressionDS)){
 for( Tissue in c('FP','TCX')){
  message(paste0("Working on: ", Tissue))
  #Tissue<-'DLPFC'

@@

Trans  <- c("ENSG00000095970", "ENSG00000198087", "ENSG00000109919", "ENSG00000074266", "ENSG00000130208", "ENSG00000103550")
names(Trans) <- c("TREM2", "CD2AP", "MTCH2", "EED", "APOC1", 'KNOP1')

@@

Cuts <- matrix(0, 101,6)
colnames( Cuts ) <- names(Partials)
row.names( Cuts ) <-  seq( from=0, to=10, by=.1 )

@@

#10% of the targets remaining at the second Standard Deviation
CutoffVal <- round( Y[18,1:6]/10 )
LIN <- rep(0, length(names(CutoffVal)))
names(LIN) <- names(CutoffVal)
for(tar in names(CutoffVal)){
  eval(parse(text=paste0( 'LIN[\'', tar, '\'] <- Y[ Y[,\'', tar, '\'] < CutoffVal[tar], ][1,\'StDevs\']')))
}

LIN <- as.data.frame(LIN)
LIN$`Seed Gene` <- row.names(LIN)
LIN$LIN <-c( 1, 1, 1, 1, 1, 1 )
LIN$LIN <- c( 1.4, 1.5, 0.4, 1.1,1.1,1.5)

@@

ggsave(g, file="Cutoffs.eps", device="eps")
system('synapse store --parentid syn21590299 --name Partial_Correlation_Decay Cutoffs.eps')
g

###Find Top 200 Targets:
REF <- c( 1.0, 1.3, 1.7, 1.7, 2.2, 2.2  )
names( REF ) <- names(NewPartials)

@@

```
```{r clusterP}
library( clusterProfiler )
library(enrichplot)

organism = "org.Hs.eg.db"
BiocManager::install(organism, character.only = TRUE)
library(organism, character.only = TRUE)
library(DOSE)

@@

  Locked_TG$Gene

@@

  Gsubset <- eval(parse(text=paste0('Target$', G )))
  Gsubset <- Gsubset[ Gsubset %in% gene.df$ENSEMBL ]
  ego <- enrichGO(gene          = Gsubset,
                universe      = gene.df$ENSEMBL,
                OrgDb         = org.Hs.eg.db,
                #keyType       = 'SYMBOL',
                keyType       = 'ENSEMBL',
                ont           = "ALL",
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.01,
                qvalueCutoff  = 0.05
              )
  if( dim(summary(ego))[1] >1 ){
    eval(parse(text=paste0( 'Enrich_PLOTS$', G, ' <- list()')))
    eval(parse(text=paste0( 'Enrich_PLOTS$', G, '$Dot <- clusterProfiler::dotplot(ego, x = "GeneRatio", showCategory=30)' )))
    eval(parse(text=paste0( 'Enrich_PLOTS$', G, '$Net <- cnetplot(ego, foldChange=NULL)' )))
    eval(parse(text=paste0( 'Enrich_PLOTS$', G, '$Erich <- enrichplot::emapplot(ego, showCategory = 10)' )))
  }else{}
}

@@

lin <- LIN
lin$LIN <- c( 0.7, 1.3,1.3,1.3,1.3,1.7 )
gG <- ggplot( data=PLOT_y, aes( x=StDevs, y=value, col=`Seed Gene`)) + 
  geom_point() + 
  geom_line() +
  geom_vline(data = lin, aes(xintercept = LIN, col=`Seed Gene`),linetype='dashed' ) +
  ylab("Partially Correlated Genes Included") +
  xlab("Standard Deviation Cutoff")

@@

setEPS()
postscript( 'LakeHeatMap.eps' )
  pheatmap::pheatmap( as.matrix(Actuals_OR), col=cols, cluster_cols=F, display_numbers = test_labels, fontsize_number=40)
dev.off()

@@

lin <- LIN
lin$LIN <- c( 0.7, 1.3,1.3,1.3,1.3,1.7 )
Mathys_G <- ggplot( data=PLOT_y, aes( x=StDevs, y=value, col=`Seed Gene`)) + 
  geom_point() + 
  geom_line() +
  geom_vline(data = lin, aes(xintercept = LIN, col=`Seed Gene`),linetype='dashed' ) +
  ylab("Partially Correlated Genes Included") +
  xlab("Standard Deviation Cutoff")

@@

Mathys_HeatMap <- pheatmap::pheatmap( as.matrix(Actuals_OR), col=cols, cluster_cols=F, display_numbers = test_labels, fontsize_number=40)

setEPS()
postscript( 'MathysHeatMap.eps' )
  pheatmap::pheatmap( as.matrix(Actuals_OR), col=cols, cluster_cols=F, display_numbers = test_labels, fontsize_number=40)
dev.off()

>>>>

Regions[ c("TREM2", "CD2AP", "MTCH2", "EED", "APOC1","CEACAM19","CLPTM1"), ]$V2 <- Regions[ c("TREM2", "CD2AP", "MTCH2", "EED", "APOC1","CEACAM19","CLPTM1"), ]$V2+500000
Regions[ c("TREM2", "CD2AP", "MTCH2", "EED", "APOC1","CEACAM19","CLPTM1"), ]$V3<-Regions[ c("TREM2", "CD2AP", "MTCH2", "EED", "APOC1","CEACAM19","CLPTM1"), ]$V3-500000

@@

#row.names(Regions) <- Regions$V4

Candidates <- list()
for( name in as.character(Regions$V4) ){

  filterlist <- paste0( Regions[ name, ]$V1, ':', Regions[ name, ]$V2, ':', Regions[ name, ]$V3 )

@@

mark <- Sys.time()
for( name in as.character(Regions$V4) ){
 #for( Tissue in names(ExpressionDS)){
 for( Tissue in names(ExpressionDS) ){
  message(paste0("Working on: ", Tissue))
  #Tissue<-'DLPFC'

@@

Trans  <- c("ENSG00000095970", "ENSG00000198087", "ENSG00000109919", "ENSG00000074266", "ENSG00000130208", "ENSG00000103550", "ENSG00000186567", "ENSG00000104853")
names(Trans) <- c("TREM2", "CD2AP", "MTCH2", "EED", "APOC1", 'KNOP1', "CEACAM19", "CLPTM1")

@@

Cuts <- matrix(0, 101,length(names(Partials)))
colnames( Cuts ) <- names(Partials)
row.names( Cuts ) <-  seq( from=0, to=10, by=.1 )

@@

#10% of the targets remaining at the second Standard Deviation
CutoffVal <- round( Y[18,1:8]/10 )
LIN <- rep(0, length(names(CutoffVal)))
names(LIN) <- names(CutoffVal)
for(tar in names(CutoffVal)){
  eval(parse(text=paste0( 'LIN[\'', tar, '\'] <- Y[ Y[,\'', tar, '\'] < CutoffVal[tar], ][1,\'StDevs\']')))
}

LIN <- as.data.frame(LIN)
LIN$`Seed Gene` <- row.names(LIN)
#LIN$LIN <-c( 1, 1, 1, 1, 1, 1 )
LIN$LIN <- c( 0.7, 1.3, 1.3, 1.3, 1.3, 1.7, 1.3, 1.3)

@@

ggsave(g, file="Cutoffs.eps", device="eps")
system('synapse store --parentid syn21818446 --name Partial_Correlation_Decay Cutoffs.eps')
g

###Find Top 200 Targets:
REF <- c( 0.7, 1.3, 1.3, 1.3, 1.3, 1.7, 1.3, 1.3 )
names( REF ) <- names(NewPartials)

@@

```
```{r clusterP}

parentId = 'syn18936948';
activityName = 'Gene Set Enrichment';
activityDescription = 'Figure 4: gene set enrichment of genes implicated in TWAS regions';
thisFileName <- 'GeneSet_Expansion.Rmd'
# Github link
thisRepo <- githubr::getRepo(repository = "jgockley62/TWAS", ref="branch", refName='master')
thisFile <- githubr::getPermlink(repository = thisRepo, repositoryPath=paste0('code/',thisFileName))

CODE <- syn_temp$store(synapseclient$Folder(name = "Gene Set Enrichment Plots", parentId = parentId))

library( clusterProfiler )
library(enrichplot)

Enrich_Run <- function( cutoffs ){
  #'@cutoffs a named vector of SD cutoff valutes for genes
  SD_Store <- rep(NA,length(names(NewPartials)))
  names(SD_Store) <- names(NewPartials)

        eval( parse( text = paste0( 'SDs <- cutoffs[\'', Gene,'\']' )))
        eval( parse( text = paste0( 'SD_Store[\'', Gene,'\'] <- ', SDs )))
        Lim <- M + SDs*S
        Adds <- colnames( Selct[ , Selct > Lim ] )
        eval(parse(text=paste0( 'Target$', Gene, ' <- c(Target$', Gene, ', Adds)'  )))
      }
    }
  }

  organism = "org.Hs.eg.db"
  #BiocManager::install(organism, character.only = TRUE)
  library(organism, character.only = TRUE)
  library(DOSE)

@@

    #Locked_TG$Gene

@@

    Gsubset <- eval(parse(text=paste0('Target$', G )))
    Gsubset <- Gsubset[ Gsubset %in% gene.df$ENSEMBL ]
    ego <- enrichGO(gene          = Gsubset,
                  universe      = gene.df$ENSEMBL,
                  OrgDb         = org.Hs.eg.db,
                  #keyType       = 'SYMBOL',
                  keyType       = 'ENSEMBL',
                  ont           = "ALL",
                  pAdjustMethod = "BH",
                  pvalueCutoff  = 0.01,
                  qvalueCutoff  = 0.05
                )

    eval(parse(text=paste0( 'SD_Used <- as.numeric(cutoffs[\'', G, '\'])' )))
    if( dim(as.data.frame(ego))[1] >1 ){
      cluster_summary <- data.frame(ego)
      write.csv(cluster_summary , file = paste0("EnrichmentResults_", G, "_Stdevs_", SD_Used,".csv"), row.names=FALSE)

      ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path=paste0("EnrichmentResults_", G, "_Stdevs_", SD_Used,".csv"), name = paste0("EnrichmentResults_", G, "_Stdevs_", SD_Used), parentId=CODE$properties$id ), activityName = activityName, executed = thisFile, activityDescription = activityDescription)

      eval(parse(text=paste0( 'Enrich_PLOTS$', G, ' <- list()')))
      # Dotplot
      eval(parse(text=paste0( 'Enrich_PLOTS$', G, '$Dot <- clusterProfiler::dotplot(ego, x = "GeneRatio", showCategory=30)' )))
      pdf(file = paste0("EnrichmentDotPlot_", G, "_Stdevs_", SD_Used,".pdf"), width = 14, height = 7)
         print( clusterProfiler::dotplot(ego, x = "GeneRatio", showCategory=30) )
      dev.off()

      ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path=paste0("EnrichmentDotPlot_", G, "_Stdevs_", SD_Used,".pdf"), name = paste0("EnrichmentDotPlot_", G, "_Stdevs_", SD_Used), parentId=CODE$properties$id ), activityName = activityName, executed = thisFile, activityDescription = activityDescription)

      # Net Plot
      eval(parse(text=paste0( 'Enrich_PLOTS$', G, '$Net <- cnetplot(ego, foldChange=NULL)' )))

      pdf(file = paste0("EnrichmentNetPlot_", G, "_Stdevs_", SD_Used,".pdf"), width = 7, height = 7)
         print( cnetplot(ego, foldChange=NULL))
      dev.off()

      ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path=paste0("EnrichmentNetPlot_", G, "_Stdevs_", SD_Used,".pdf"), name = paste0("EnrichmentNetPlot_", G, "_Stdevs_", SD_Used), parentId=CODE$properties$id ), activityName = activityName, executed = thisFile, activityDescription = activityDescription)

      # Enrichment Plot
      eval(parse(text=paste0( 'Enrich_PLOTS$', G, '$Erich <- enrichplot::emapplot(ego, showCategory = 10)' )))

      pdf(file = paste0("EnrichmentEnrichPlot_", G, "_Stdevs_", SD_Used,".pdf"), width = 7, height = 7)
        print( enrichplot::emapplot(ego, showCategory = 10) )
      dev.off()

      ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path=paste0("EnrichmentEnrichPlot_", G, "_Stdevs_", SD_Used,".pdf"), name = paste0("EnrichmentEnrichPlot_", G, "_Stdevs_", SD_Used,".pdf"), parentId=CODE$properties$id ), activityName = activityName, executed = thisFile, activityDescription = activityDescription)

    }else{}
  }
}

#Run Enrichment at Broad Cutoffs:
input <- c( TREM2 = 0.8, 
            CD2AP= 0.8, 
            MTCH2= 0.8, 
            EED= 0.8, 
            KNOP1= 0.8, 
            APOC1= 0.8,
            CEACAM19= 0.8,
            CLPTM1= 0.8
          )
Enrich_Run( input )

#Run Enrichment at Narrow Cutoffs:
input <- c( TREM2 = 2.8, 
            CD2AP = 2.8, 
            MTCH2 = 2.8, 
            EED = 2.8, 
            KNOP1 = 2.8, 
            APOC1 = 2.8,
            CEACAM19= 2.8,
            CLPTM1= 2.8
          )
Enrich_Run( input )

#Run Enrichment at Custom Cutoffs:
input <- c( TREM2 = 0.8, 
            #CD2AP = 2, 
            CD2AP = 1,
            MTCH2 = 0.4, 
            EED = 2, 
            KNOP1 = 2, 
            APOC1 = 1.7,
            CEACAM19= 2,
            CLPTM1= 2
          )
Enrich_Run( input )
#if( Gene %in% c( "TREM2")){
          #SDs <- 0.7
          #SDs <- 2.8
        #}else{
          #if( Gene %in% c( 'APOC1')){
            #SDs <- 1.7
            #SDs <- 2.8
          #}else{
            #SDs <- 1.3
            #SDs <- 2.8
          #}
        #}
        #if( Gene %in% c( "EED")){
          #SDs <- 0.5
        #}else{
          #if( Gene %in% c( "KNOP1")){
           # SDs <- 0.8
          #}else{
            #if( Gene %in% c( 'MTCH2') ){
              #SDs <- 0.5
              #SDs <- 0.4
              #SDs <- 1.5
            #}else{
              #if( Gene %in% c( 'TREM2')){
                #SDs <- 1.4
              #}else{
                #SDs <- 1.5
              #}
            #}
          #}
        #}

@@

lin <- LIN
lin$LIN <- c( 0.7, 1.3,1.3,1.3,1.3,1.7,1.3,1.3 )
gG <- ggplot( data=PLOT_y, aes( x=StDevs, y=value, col=`Seed Gene`)) + 
  geom_point() + 
  geom_line() +
  geom_vline(data = lin, aes(xintercept = LIN, col=`Seed Gene`),linetype='dashed' ) +
  ylab("Partially Correlated Genes Included") +
  xlab("Standard Deviation Cutoff")

@@

save_pheatmap_pdf <- function(x, filename, width=7, height=7) {
   stopifnot(!missing(x))
   stopifnot(!missing(filename))
   pdf(filename, width=width, height=height)
   grid::grid.newpage()
   grid::grid.draw(x$gtable)
   dev.off()
}

save_pheatmap_pdf(Lake_HeatMap, "Lake_HeatMap.pdf")

@@

lin <- LIN
lin$LIN <- c( 0.7, 1.3,1.3,1.3,1.3,1.7,1.3,1.3 )
Mathys_G <- ggplot( data=PLOT_y, aes( x=StDevs, y=value, col=`Seed Gene`)) + 
  geom_point() + 
  geom_line() +
  geom_vline(data = lin, aes(xintercept = LIN, col=`Seed Gene`),linetype='dashed' ) +
  ylab("Partially Correlated Genes Included") +
  xlab("Standard Deviation Cutoff")

@@

Mathys_HeatMap <- pheatmap::pheatmap( as.matrix(Actuals_OR), col=cols, cluster_cols=F, display_numbers = test_labels, fontsize_number=40)
save_pheatmap_pdf(Mathys_HeatMap, "Mathys_HeatMap.pdf")

parentId = 'syn18936948';
activityName = 'Gene Set Enrichment';
activityDescription = 'Figure 4: gene set enrichment of genes implicated in TWAS regions';
thisFileName <- 'GeneSet_Expansion.Rmd'
# Github link
thisRepo <- githubr::getRepo(repository = "jgockley62/TWAS", ref="branch", refName='master')
thisFile <- githubr::getPermlink(repository = thisRepo, repositoryPath=paste0('code/',thisFileName))

CODE <- syn_temp$store(synapseclient$Folder(name = "Gene Set Enrichment Plots", parentId = parentId))

ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='STDEV_Cutoff.eps', name = 'StDev Cutoff values of Partial Correlations', parentId=CODE$properties$id ), activityName = activityName, executed = thisFile, activityDescription = activityDescription)

ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='Lake_HeatMap.pdf', name = 'Lake et.al CellType Specificity', parentId=CODE$properties$id ), activityName = activityName, executed = thisFile, activityDescription = activityDescription)

ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='Mathys_HeatMap.pdf', name = 'Mathys et.al CellType Specificity', parentId=CODE$properties$id ), activityName = activityName, executed = thisFile, activityDescription = activityDescription)
